---
title: "Housing Price- New V1.0"
author: "Sumit Dhar"
date: "July 6, 2016"
output: html_document
---
Using the Seattle area King County ~ 26K data set to do something hopefully interesting!!

```{r}

dir <- 'D:\\MyStuff\\Housing Price'
setwd(dir)
sales <- read.csv('kc_house_data.csv', stringsAsFactors = FALSE)
#Back up data set
bsales <- sales

#str(sales)
```
#Initial Manipulation
Looking at the data to create initial features for modeling

```{r}
sales$saleyrmnth <- as.integer(format(substr(sales$date,1,6), format ="Y%-m%"))
unique(sales$saleyrmnth)
sales$sdate <- as.Date(paste0(
  format(substr(sales$date,1,4), format = "Y%-m%-d%"),'-',format(substr(sales$date,5,6), format =
                                                                   "Y%-m%-d%"),'-',format(substr(sales$date,7,8), format =
                                                                                            "Y%-m%-d%")
))
# Total Sqft
sales$totsqft <- sales$sqft_living + sales$sqft_lot
# Above sqft to total
sales$sqftabvpct <- sales$sqft_above/(sales$sqft_above+sales$sqft_basement)
# Proportion of bedrooms
sales$bedprop <- sales$bedrooms/(sales$bedrooms+sales$bathrooms+1)
#Age of property
sales$age <- sales$age <-as.numeric(substr(sales$date,1,4))- sales$yr_built
#For building with no renovation imputing year of contruction
sales[sales$yr_renovated==0,"yr_renovated"] <- sales$yr_built[sales$yr_renovated==0]
#Defining new variable Reage
sales$reage <-as.numeric(substr(sales$date,1,4))- sales$yr_renovated
#Ratiof living to lot
sales$liv2lot <- sales$sqft_living/sales$sqft_lot
#"grade" 
class(sales$grade)
sales$grade1 <- as.character(sales$grade)
sales$grade1 <- 'Between 6 &10'
sales$grade1[sales$grade > 10] <- '>10'
sales$grade1[sales$grade < 6] <- '<6'
#str(sales)
```
Exploring zip code wise
```{r}
aggdata <- aggregate(
    sales$price ~ sales$zipcode, data = sales, FUN = function(x) {
        median(x)
    }
)
names(aggdata) <- c('zipcode','medianprice')
library(gtools)
aggdata$zipdec <-quantcut( aggdata$medianprice, seq(0,1,by=0.1) )
for (i in 1:nrow(aggdata)){
  sales$zipdec[sales$zipcode == aggdata[i,1]] <- aggdata[i,3]
}
```
We shall take zipcode distance from 98039, 98004, 98040, 98112
```{r}
sales$zipd_98039 <- sales$zipcode- 98039
sales$zipd_98004 <- sales$zipcode- 98004
sales$zipd_98040 <- sales$zipcode- 98040
sales$zipd_98112 <- sales$zipcode- 98112
```
# Some visualization
```{r}
library(ggplot2)
ggplot(sales, aes(as.factor(grade1),price)) +
  geom_boxplot(color="blue", fill="green") +
  ggtitle("Box Plot for Building Grade") +
  xlab("Grade")+ ylab("Price")

ggplot(sales, aes(as.factor(sales$saleyrmnth), fill=sales$price)) + geom_bar()+
  xlab("Month")+ ylab("Price($)")
```
Clearly shows seasonality
```{r}
ggplot(sales, aes(as.factor(bedrooms^2),price)) +
  geom_boxplot(color="blue", fill="green") +
  ggtitle("Box Plot for Bedrooms in Seattle") +
  xlab("Bedrooms^2") + ylab("Price($)")
```

```{r}
ggplot(sales, aes(as.factor(bedrooms*bathrooms),price)) +
  geom_boxplot(color="blue", fill="green") +
  ggtitle("Box Plot for Bedroomsxbathrooms in Seattle") +
  xlab("BedroomsxBathrooms") + ylab("Price($)")
```

```{r}
ggplot(sales, aes(as.factor(bathrooms),price)) +
  geom_boxplot(color="blue", fill="green") +
  ggtitle("Box Plot for Bathrooms in Seattle") +
  xlab("Bathrooms") + ylab("Price")
```
Number of Bathrooms has a definite impact
```{r}
ggplot(sales, aes(as.factor(bedprop),price)) +
  geom_boxplot(color="blue", fill="green") +
  ggtitle("Box Plot for Bathrooms Prop. in Seattle") +
  xlab("Bathrooms :Prop.") + ylab("Price")
```

Hump in the middle
```{r}
##"floors"
ggplot(sales, aes(as.factor(floors),price)) +
  geom_boxplot(color="blue", fill="green") +
  ggtitle("Box Plot for Floors in Seattle") +
  xlab("Floors") + ylab("Price")
```

```{r}
qplot(sales$sqft_living/sales$floors,sales$price,colour= sales$price,
      xlab ="Square feet/ Floor", ylab="Price($)")
```


```{r}
library(gtools)
sales$psqftdec <-quantcut( sales$price/sales$totsqft, seq(0,1,by=0.1) )
sales$pricedec <-quantcut( sales$price, seq(0,1,by=0.1) )
qplot(sales$zipd_98039,sales$price/sales$totsqft,colour= sales$psqftdec,
      xlab ="ZipCode Distance from 98039", ylab="Psqft($)")
```

```{r}
#[9] "waterfront"   
ggplot(sales, aes(as.factor(waterfront),price)) +
  geom_boxplot(color="blue", fill="green") +
  ggtitle("Box Plot for Bedrooms in Seattle") +
  xlab("Waterfront") + ylab("Price")
```

## Spliting Training and Test Data

```{r}
sales1 <- sales
sales <-
  subset(sales, select = -c(id, date, yr_built, yr_renovated, pricedec, grade, saleyrmnth, zipcode))
lapply(sales, function(x){sum(!is.finite(x))})
lapply(sales, function(x){sum(is.na(x))})
lapply(sales, function(x){sum(is.nan(x))})
```
All variables appear finite. This is required for ensuring that the various algorithms work correctly
```{r}
set.seed(12345)
smp_size <- 0.8*nrow(sales)
train_ind <- sample(seq_len(nrow(sales)), size = smp_size)

train <- sales[train_ind, ]
test <- sales[-train_ind, ]
#str(sales)
```
#Single variable regression
```{r}
coln <- names(train)
rsq <- seq(1:ncol(train))
voe<- seq(1:ncol(train))
rsq<-seq(1:ncol(train)) 
intr<- seq(1:ncol(train))
for (i in 1:ncol(train)) {
  Formula <- paste0("train$price~",coln[i])
  lm(Formula, train)
    sing.var.reg<-lm(Formula, train)
    rsq[i]<- summary(sing.var.reg)$r.squared
    voe[i]<-sqrt(deviance(sing.var.reg)/df.residual(sing.var.reg))
    intr[i]<- coefficients(sing.var.reg)[1]
  }
sin.var.result <- cbind(coln,rsq,voe,intr)
print(sin.var.result[order(-rsq,-voe),])
```
Trying to understand interaction between bath and bed
```{r}
sin.var.reg.v2 <- lm(price~poly(bedrooms*bathrooms,2),train)
summary(sin.var.reg.v2)
```

#First Multiple regression
```{r}
library(glmnet)
Mul.var.reg.v1 <- glm(price~ sqft_living+
                       bedrooms+
                       bathrooms+
                       sqft_living+
                       sqft_lot+
                       as.factor(floors)+
                       as.factor(waterfront)+
                       as.factor(view)+
                       as.factor(condition)+
                       as.factor(grade1)+
                       as.factor(zipdec)+
                       sqft_above+
                       sqft_basement+
                       poly(lat*long,2)+
                       sqft_living15+
                       sqft_lot15+
                       sdate+
                       totsqft+
                       sqftabvpct+
                       bedprop+
                       age+
                       reage+
                       liv2lot+
                       zipd_98112+
                       zipd_98039+
                       zipd_98004+
                       zipd_98112, train, family=gaussian())

summary(Mul.var.reg.v1)
test$pred <- predict(Mul.var.reg.v1, test)
sqrt(mean((test$price - test$pred)^2))

qplot(test$price,test$pred,colour= (test$price - test$pred)^2,
      xlab ="Actual($)", ylab="Predicted($)")+
  geom_smooth(method = "lm")
```
# 2nd Multiple Regression
```{r}
Mul.var.reg.v2 <- glm(price~ sqft_living+
                       bedrooms*bathrooms+
                       sqft_living+
                       sqft_lot+
                       as.factor(floors)+
                       as.factor(waterfront)+
                       as.factor(view)+
                       as.factor(condition)+
                       as.factor(grade1)+
                       sqft_above+
                       sqft_basement+
                       as.factor(zipdec)+
                       poly(lat*long,2)+
                       sqft_living15+
                       sqft_lot15+
                       sdate+
                       sqftabvpct+
                       bedprop+
                       age+
                       reage+
                       liv2lot, train, family=gaussian())

summary(Mul.var.reg.v2)
test$pred <- predict(Mul.var.reg.v2, test)
sqrt(mean((test$price - test$pred)^2))

qplot(test$price,test$pred,colour= (test$price - test$pred)^2,
      xlab ="Actual($)", ylab="Predicted($)")+
  geom_smooth(method = "lm")

```
#3rd Linear Regression
```{r}
Mul.var.reg.v3 <- glm(price~ sqft_living+
                       bedrooms*bathrooms+
                       sqft_living+
                       sqft_lot+
                       sdate+
                       as.factor(floors)+
                       as.factor(waterfront)+
                       as.factor(view)+
                       as.factor(condition)+
                       as.factor(grade1)+
                       sqft_above+
                       as.factor(zipdec)+
                       poly(lat*long,2)+
                       sqft_living15+
                       sqftabvpct+
                       bedprop+
                       age+
                       reage+
                       liv2lot, train, family=gaussian())

summary(Mul.var.reg.v3)
test$pred <- predict(Mul.var.reg.v3, test)
sqrt(mean((test$price - test$pred)^2))

qplot(test$price,test$pred,colour= (test$price - test$pred)^2,
      xlab ="Actual($)", ylab="Predicted($)")+
  geom_smooth(method = "lm")

```

# 10 Fold cross validation
```{r}
library(boot)
k= 10
cv.error.10=rep(0,k)
for (i in 1:k){
    cv.error.10[i]=cv.glm(train,Mul.var.reg.v3,K=10)$delta[1]
}
print(paste0("K=",k," times Cross Validation Error: ",sqrt(sum(cv.error.10)/k)))
```
# LASSO Regression
```{r}
grid = 10 ^ seq(10,-2,length = 100)
x = model.matrix(price ~ sqft_living+
                       bedrooms+
                       bathrooms+
                       sqft_living+
                       sdate+
                       sqft_lot+
                       as.factor(floors)+
                       as.factor(waterfront)+
                       as.factor(view)+
                       as.factor(condition)+
                       as.factor(grade1)+
                       sqft_above+
                       sqft_basement+
                       as.factor(zipdec)+
                       poly(lat*long,2)+
                       sqft_living15+
                       sqft_lot15+
                       totsqft+
                       sqftabvpct+
                       bedprop+
                       age+
                       reage+
                       liv2lot+
                       zipd_98112+
                       zipd_98039+
                       zipd_98004+
                       zipd_98112,sales)
y = sales$price
lasso.mod.v1 = glmnet(x[train_ind,],y[train_ind],alpha = 1,lambda = grid)
plot(lasso.mod.v1)
cv.out = cv.glmnet(x[train_ind,],y[train_ind],alpha = 1, k=10)
plot(cv.out)
bestlam = cv.out$lambda.min
print(bestlam)
lasso.pred = predict(lasso.mod.v1,s = bestlam,newx = x[-train_ind,])
print(sqrt(mean((lasso.pred - y[-train_ind]) ^ 2)))
out = glmnet(x,y,alpha = 1,lambda = grid)
lasso.coef = predict(out,type = "coefficients",s = bestlam)[1:20,]
lasso.coef
lasso.coef[lasso.coef != 0]
qplot(y[-train_ind], lasso.pred,colour=(lasso.pred - y[-train_ind]) ^ 2,
      xlab ="Actual($)", ylab="Predicted($)")+
  geom_smooth(method = "lm")+
  ggtitle("LASSO Regression Pred vs. Actual in Test")
```

```{r}
x = model.matrix(price ~ sqft_living+
                       bedrooms+
                       bathrooms+
                       sqft_living+
                       sdate+
                       sqft_lot+
                       floors+
                       waterfront+
                       view+
                       condition+
                       grade1+
                       sqft_above+
                       sqft_basement+
                       as.factor(zipdec)+
                       poly(lat*long,2)+
                       sqft_living15+
                       sqft_lot15+
                       totsqft+
                       sqftabvpct+
                       bedprop+
                       age+
                       reage+
                       liv2lot+
                       zipd_98112+
                       zipd_98039+
                       zipd_98004+
                       zipd_98112,sales)[,-1]
y = sales$price
lasso.mod.v2 = glmnet(x[train_ind,],y[train_ind],alpha = 1,lambda = grid)
plot(lasso.mod.v2)
cv.out = cv.glmnet(x[train_ind,],y[train_ind],alpha = 1, k=10)
plot(cv.out)
bestlam = cv.out$lambda.min
print(bestlam)
lasso.pred = predict(lasso.mod.v2,s = bestlam,newx = x[-train_ind,])
print(sqrt(mean((lasso.pred - y[-train_ind]) ^ 2)))
out = glmnet(x,y,alpha = 1,lambda = grid)
lasso.coef = predict(out,type = "coefficients",s = bestlam)[1:20,]
lasso.coef
lasso.coef[lasso.coef != 0]
qplot(y[-train_ind], lasso.pred,colour=(lasso.pred - y[-train_ind]) ^ 2,
      xlab ="Actual($)", ylab="Predicted($)")+
  geom_smooth(method = "lm")+
  ggtitle("LASSO Regression Pred vs. Actual in Test")
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```